---
title: "Descriptive Data"
format: bookup-html
---

```{r setup, include=FALSE}
# Load required libraries
library(leaflet)
library(tidyverse)
library(viridis)
library(ggplot2)
library(dplyr)
library(plotly)
library(sf)
library(htmltools)

```

```{r data-import, include=FALSE}
# Import spatial data
library(sf)
model_data_hpi <- st_read("data/model_data_hpi.shp")
```


# Spatial Distribution of Utah's Correctional Population

```{r spatial, echo=FALSE}
# Assuming model_data_hpi is your dataset
leaflet(model_data_hpi) %>%
  addTiles() %>%
  addPolygons(
    fillColor = ~colorQuantile("YlOrRd", corr_rt)(corr_rt),
    fillOpacity = 0.7,
    weight = 1,
    color = "#FFFFFF",
    popup = ~paste(
      "<strong>Tract:</strong>", NAM, "<br>",
      "<strong>Rate:</strong>", round(corr_rt, 2), "<br>",
      "<strong>Hispanic:</strong>", round(phisp, 2), "<br>",
      "<strong>Unemployment:</strong>", round(punemp, 2)
    )
  ) %>%
  addLegend(
    position = "bottomright",
    pal = colorQuantile("YlOrRd", model_data_hpi$corr_rt),
    values = model_data_hpi$corr_rt,
    title = "Correctional Population Rate </br>per 1000 individuals"
  )
```

# Correctional Population Outlier Analysis

This section presents outlier analysis for correctional population counts (`corr_pop`) and rates (`corr_rate`) using different outlier treatment approaches.

This analysis examines correctional population patterns using:

- **Population Counts (`corr_pop`)**: Raw counts of individuals under supervision in each census tract

- **Population Rates (`corr_rate`)**: Rates per 1,000 population for standardized comparison in each census tractw

::: {.panel-tabset}
## Distribution Histogram: Correctional Population
```{r}
#| echo: false
#| results: asis
htmltools::tags$iframe(
  src = "descriptive/corr_pop_histogram.html",
  width = "100%",
  height = "600px",
  frameborder = "0",
  style = "border: none;"
)

```

## Distribution Histogram: Correctional Rate
```{r}
#| echo: false
#| results: asis
htmltools::tags$iframe(
  src = "descriptive/corr_rate_histogram.html",
  width = "100%",
  height = "600px",
  frameborder = "0",
  style = "border: none;"
)
```

:::

# Distribution Tables

- **High Outliers (95th Percentile) Correctional Population**: 196.8

- **High Outliers (95th Percentile) Correctional Rate**: 45.69

- **Low Outliers (5th Percentile) Correctional Population**: 11

- **Low Outliers (5th Percentile) Correctional Rate**: 2.73


## Summary Table
```{r}
#| echo: false
#| results: asis
htmltools::tags$iframe(
  src = "descriptive/summary_table_flextable.html",
  width = "100%",
  height = "600px", 
  frameborder = "0"
)

```


<br>

::: {.panel-tabset}

## Population Counts 🔢 - Low Outlier Treatment
```{r}
#| echo: false
#| results: asis

#htmltools::includeHTML("descriptive/corr_pop_LOW_outliers_table.html")

htmltools::div(
  # Inline table (smaller)
  htmltools::div(
    style = "position: relative; border: 1px solid #ddd; margin: 1rem 0;",
    htmltools::tags$button(
      onclick = "openModal('table-modal')",
      style = "position: absolute; top: 5px; right: 5px; z-index: 100; background: #28a745; color: white; border: none; padding: 8px 15px; border-radius: 4px; cursor: pointer;",
      "🔍 View Full Size"
    ),
    htmltools::tags$iframe(
      src = "descriptive/corr_pop_LOW_outliers_table.html",
      width = "100%",
      height = "400px",
      frameborder = "0",
      style = "transform: scale(0.8); transform-origin: top left; width: 125%; height: 500px;"
    )
  ),
  
  # Modal overlay
  htmltools::div(
    id = "table-modal",
    style = "display: none; position: fixed; z-index: 9999; left: 0; top: 0; width: 100%; height: 100%; background-color: rgba(0,0,0,0.5);",
    htmltools::div(
      style = "position: relative; margin: 2% auto; width: 95%; height: 90%; background: white; border-radius: 8px; padding: 20px;",
      htmltools::tags$button(
        onclick = "closeModal('table-modal')",
        style = "position: absolute; top: 10px; right: 15px; background: #dc3545; color: white; border: none; padding: 8px 15px; border-radius: 4px; cursor: pointer; font-weight: bold;",
        "✕ Close"
      ),
      htmltools::tags$iframe(
        src = "descriptive/corr_pop_LOW_outliers_table.html",
        width = "100%",
        height = "95%",
        frameborder = "0"
      )
    )
  ),
  
  htmltools::tags$script(HTML("
    function openModal(id) {
      document.getElementById(id).style.display = 'block';
      document.body.style.overflow = 'hidden';
    }
    
    function closeModal(id) {
      document.getElementById(id).style.display = 'none';
      document.body.style.overflow = 'auto';
    }
    
    // Close modal when clicking outside
    window.onclick = function(event) {
      const modal = document.getElementById('table-modal');
      if (event.target === modal) {
        closeModal('table-modal');
      }
    }
  "))
)
```

## Population Counts 🔢 - High Outlier Treatment
```{r}
#| echo: false
#| results: asis
#htmltools::includeHTML("descriptive/corr_pop_HIGH_outliers_table.html")

htmltools::div(
  # Inline table (smaller)
  htmltools::div(
    style = "position: relative; border: 1px solid #ddd; margin: 1rem 0;",
    htmltools::tags$button(
      onclick = "openModal('table-modal')",
      style = "position: absolute; top: 5px; right: 5px; z-index: 100; background: #28a745; color: white; border: none; padding: 8px 15px; border-radius: 4px; cursor: pointer;",
      "🔍 View Full Size"
    ),
    htmltools::tags$iframe(
      src = "descriptive/corr_pop_HIGH_outliers_table.html",
      width = "100%",
      height = "400px",
      frameborder = "0",
      style = "transform: scale(0.8); transform-origin: top left; width: 125%; height: 500px;"
    )
  ),
  
  # Modal overlay
  htmltools::div(
    id = "table-modal",
    style = "display: none; position: fixed; z-index: 9999; left: 0; top: 0; width: 100%; height: 100%; background-color: rgba(0,0,0,0.5);",
    htmltools::div(
      style = "position: relative; margin: 2% auto; width: 95%; height: 90%; background: white; border-radius: 8px; padding: 20px;",
      htmltools::tags$button(
        onclick = "closeModal('table-modal')",
        style = "position: absolute; top: 10px; right: 15px; background: #dc3545; color: white; border: none; padding: 8px 15px; border-radius: 4px; cursor: pointer; font-weight: bold;",
        "✕ Close"
      ),
      htmltools::tags$iframe(
        src = "descriptive/corr_pop_HIGH_outliers_table.html",
        width = "100%",
        height = "95%",
        frameborder = "0"
      )
    )
  ),
  
  htmltools::tags$script(HTML("
    function openModal(id) {
      document.getElementById(id).style.display = 'block';
      document.body.style.overflow = 'hidden';
    }
    
    function closeModal(id) {
      document.getElementById(id).style.display = 'none';
      document.body.style.overflow = 'auto';
    }
    
    // Close modal when clicking outside
    window.onclick = function(event) {
      const modal = document.getElementById('table-modal');
      if (event.target === modal) {
        closeModal('table-modal');
      }
    }
  "))
)

```

## Population Rates 📊 - Low Outlier Treatment
```{r}
#| echo: false
#| results: asis
#htmltools::includeHTML("descriptive/corr_rate_LOW_outliers_table.html")

htmltools::div(
  # Inline table (smaller)
  htmltools::div(
    style = "position: relative; border: 1px solid #ddd; margin: 1rem 0;",
    htmltools::tags$button(
      onclick = "openModal('table-modal')",
      style = "position: absolute; top: 5px; right: 5px; z-index: 100; background: #28a745; color: white; border: none; padding: 8px 15px; border-radius: 4px; cursor: pointer;",
      "🔍 View Full Size"
    ),
    htmltools::tags$iframe(
      src = "descriptive/corr_rate_LOW_outliers_table.html",
      width = "100%",
      height = "400px",
      frameborder = "0",
      style = "transform: scale(0.8); transform-origin: top left; width: 125%; height: 500px;"
    )
  ),
  
  # Modal overlay
  htmltools::div(
    id = "table-modal",
    style = "display: none; position: fixed; z-index: 9999; left: 0; top: 0; width: 100%; height: 100%; background-color: rgba(0,0,0,0.5);",
    htmltools::div(
      style = "position: relative; margin: 2% auto; width: 95%; height: 90%; background: white; border-radius: 8px; padding: 20px;",
      htmltools::tags$button(
        onclick = "closeModal('table-modal')",
        style = "position: absolute; top: 10px; right: 15px; background: #dc3545; color: white; border: none; padding: 8px 15px; border-radius: 4px; cursor: pointer; font-weight: bold;",
        "✕ Close"
      ),
      htmltools::tags$iframe(
        src = "descriptive/corr_rate_LOW_outliers_table.html",
        width = "100%",
        height = "95%",
        frameborder = "0"
      )
    )
  ),
  
  htmltools::tags$script(HTML("
    function openModal(id) {
      document.getElementById(id).style.display = 'block';
      document.body.style.overflow = 'hidden';
    }
    
    function closeModal(id) {
      document.getElementById(id).style.display = 'none';
      document.body.style.overflow = 'auto';
    }
    
    // Close modal when clicking outside
    window.onclick = function(event) {
      const modal = document.getElementById('table-modal');
      if (event.target === modal) {
        closeModal('table-modal');
      }
    }
  "))
)

```

## Population Rates 📊 - High Outlier Treatment
```{r}
#| echo: false
#| results: asis
#htmltools::includeHTML("descriptive/corr_rate_HIGH_outliers_table.html")

htmltools::div(
  # Inline table (smaller)
  htmltools::div(
    style = "position: relative; border: 1px solid #ddd; margin: 1rem 0;",
    htmltools::tags$button(
      onclick = "openModal('table-modal')",
      style = "position: absolute; top: 5px; right: 5px; z-index: 100; background: #28a745; color: white; border: none; padding: 8px 15px; border-radius: 4px; cursor: pointer;",
      "🔍 View Full Size"
    ),
    htmltools::tags$iframe(
      src = "descriptive/corr_rate_HIGH_outliers_table.html",
      width = "100%",
      height = "400px",
      frameborder = "0",
      style = "transform: scale(0.8); transform-origin: top left; width: 125%; height: 500px;"
    )
  ),
  
  # Modal overlay
  htmltools::div(
    id = "table-modal",
    style = "display: none; position: fixed; z-index: 9999; left: 0; top: 0; width: 100%; height: 100%; background-color: rgba(0,0,0,0.5);",
    htmltools::div(
      style = "position: relative; margin: 2% auto; width: 95%; height: 90%; background: white; border-radius: 8px; padding: 20px;",
      htmltools::tags$button(
        onclick = "closeModal('table-modal')",
        style = "position: absolute; top: 10px; right: 15px; background: #dc3545; color: white; border: none; padding: 8px 15px; border-radius: 4px; cursor: pointer; font-weight: bold;",
        "✕ Close"
      ),
      htmltools::tags$iframe(
        src = "descriptive/corr_pop_HIGH_outliers_table.html",
        width = "100%",
        height = "95%",
        frameborder = "0"
      )
    )
  ),
  
  htmltools::tags$script(HTML("
    function openModal(id) {
      document.getElementById(id).style.display = 'block';
      document.body.style.overflow = 'hidden';
    }
    
    function closeModal(id) {
      document.getElementById(id).style.display = 'none';
      document.body.style.overflow = 'auto';
    }
    
    // Close modal when clicking outside
    window.onclick = function(event) {
      const modal = document.getElementById('table-modal');
      if (event.target === modal) {
        closeModal('table-modal');
      }
    }
  "))
)

```

:::

## Combined Tables
- **Both Population & Rate Tables** are presented below including both HIGH and LOW outliers with flags for treatment center in census tract 

```{r}
#| echo: false
#| results: asis
#htmltools::includeHTML("descriptive/corr_pop_outliers_table.html")

htmltools::div(
  # Inline table (smaller)
  htmltools::div(
    style = "position: relative; border: 1px solid #ddd; margin: 1rem 0;",
    htmltools::tags$button(
      onclick = "openModal('table-modal')",
      style = "position: absolute; top: 5px; right: 5px; z-index: 100; background: #28a745; color: white; border: none; padding: 8px 15px; border-radius: 4px; cursor: pointer;",
      "🔍 View Full Size"
    ),
    htmltools::tags$iframe(
      src = "descriptive/corr_pop_outliers_table.html",
      width = "100%",
      height = "400px",
      frameborder = "0",
      style = "transform: scale(0.8); transform-origin: top left; width: 125%; height: 500px;"
    )
  ),
  
  # Modal overlay
  htmltools::div(
    id = "table-modal",
    style = "display: none; position: fixed; z-index: 9999; left: 0; top: 0; width: 100%; height: 100%; background-color: rgba(0,0,0,0.5);",
    htmltools::div(
      style = "position: relative; margin: 2% auto; width: 95%; height: 90%; background: white; border-radius: 8px; padding: 20px;",
      htmltools::tags$button(
        onclick = "closeModal('table-modal')",
        style = "position: absolute; top: 10px; right: 15px; background: #dc3545; color: white; border: none; padding: 8px 15px; border-radius: 4px; cursor: pointer; font-weight: bold;",
        "✕ Close"
      ),
      htmltools::tags$iframe(
        src = "descriptive/corr_pop_outliers_table.html",
        width = "100%",
        height = "95%",
        frameborder = "0"
      )
    )
  ),
  
  htmltools::tags$script(HTML("
    function openModal(id) {
      document.getElementById(id).style.display = 'block';
      document.body.style.overflow = 'hidden';
    }
    
    function closeModal(id) {
      document.getElementById(id).style.display = 'none';
      document.body.style.overflow = 'auto';
    }
    
    // Close modal when clicking outside
    window.onclick = function(event) {
      const modal = document.getElementById('table-modal');
      if (event.target === modal) {
        closeModal('table-modal');
      }
    }
  "))
)

```


