---
title: "Policy Applications and Recommendations"
format: bookup-html
---
```{r inla-setup}
library(tidyverse)
library(sf)
library(tidycensus)
library(tigris)
library(spdep)
library(INLA)
library(BMA)
library(leaflet)
library(plotly)
library(DT)
library(kableExtra)
library(corrplot)
library(car)
library(viridis)
library(patchwork)

# Set global options
options(tigris_use_cache = TRUE)
sf_use_s2(FALSE)

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  cache = TRUE
)


# Your INLA setup code here
```

# Translating Research to Policy

This chapter demonstrates how spatial analysis informs evidence-based policy...

# Part VI: Policy Translation Framework

## 6.1 Evidence-Based Policy Targeting

### Priority Matrix for Intervention

```{r policy-matrix}
policy_interventions <- data.frame(
  Priority = 1:8,
  `Policy Area` = c("Environmental Justice", "Transportation Access", "Education Access", 
                    "Community Engagement", "Health Insurance", "Housing Quality",
                    "Employment Programs", "Economic Development"),
  `Effect Size` = c("+494.7%", "-91.0%", "-88.4%", "-73.6%", "-73.3%", 
                    "+59.9%", "-50.0%", "+49.9%"),
  `Evidence Level` = c("Strong", "Strong", "Strong", "Strong", "Strong",
                      "Moderate", "Moderate", "Suggestive"),
  `Implementation Strategy` = c(
    "Environmental monitoring and remediation in high-risk areas",
    "Transit development and auto access programs", 
    "Educational opportunity expansion in underserved areas",
    "Community organizing and civic engagement initiatives",
    "Healthcare access expansion and insurance enrollment",
    "Housing rehabilitation and quality improvement programs",
    "Job training and placement in high-need areas",
    "Economic development targeting capability enhancement"
  )
)

DT::datatable(policy_interventions,
              caption = "Policy Priority Matrix: Evidence-Based Intervention Targeting",
              options = list(pageLength = 8, dom = 't'),
              class = 'stripe hover') %>%
  formatStyle("Effect.Size", 
              backgroundColor = styleInterval(c(0), c("#ffebee", "#e8f5e8")))
```

## 6.2 Geographic Targeting Strategy

### Using Spatial Results for Resource Allocation

::: {.panel-tabset}

## High-Risk Areas (Negative u_i)

**Characteristics**: Areas with worse-than-expected outcomes despite measured characteristics

**Strategy**: **Intensive Investment and Capacity Building**
- Expand treatment and rehabilitation options
- Increase job training and employment programs  
- Improve transportation infrastructure
- Build community support systems
- Address environmental hazards
- Enhance service coordination

## High-Capability Areas (Positive u_i) 

**Characteristics**: Areas with better-than-expected outcomes

**Strategy**: **Asset-Based Development and Replication**
- Build on existing community strengths
- Enhance coordination and expand successful programs
- Share resources and expertise with other areas
- Document and replicate effective practices
- Support regional leadership development

## Transition Areas (Low |u_i|)

**Characteristics**: Outcomes well-explained by measured individual factors

**Strategy**: **Individual-Focused Services**
- Emphasize targeted individual support services
- Address specific personal and family barriers
- Provide direct service delivery
- Focus on individual skill development and support

:::

## 6.3 Implementation Framework

### Multi-Level Intervention Strategy

```{mermaid}
flowchart TD
    A[INLA Spatial Analysis Results] --> B[Geographic Targeting]
    A --> C[Factor Prioritization]
    
    B --> D[High-Risk Areas<br/>Intensive Investment]
    B --> E[High-Capability Areas<br/>Asset Building]
    B --> F[Individual-Focus Areas<br/>Direct Services]
    
    C --> G[Environmental Justice<br/>Air Quality Improvement]
    C --> H[Transportation Access<br/>Mobility Enhancement]  
    C --> I[Community Engagement<br/>Collective Efficacy]
    
    D --> J[Policy Implementation]
    E --> J
    F --> J
    G --> J
    H --> J
    I --> J
    
    J --> K[Monitoring & Evaluation]
    K --> L[Adaptive Management]
    L --> A
    
    style A fill:#e1f5fe
    style J fill:#f3e5f5
    style K fill:#e8f5e8
```

